---
title: "Text Mining of Taylor Swift Lyrics"
author: "Carson Zhang (cxzhang4)"
date: "3/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(tidytext)
library(gridExtra)

library(dplyr)
library(ggplot2)
library(purrr)
library(tibble)
library(tidyr)
library(broom)
```

```{r}
ts_lyrics = read_csv("data/taylor_swift_lyrics.csv")
ts_lyrics_tidy = ts_lyrics %>% unnest_tokens(word, lyric)
```

```{r}
# albums = unique(ts_lyrics$album)
# album_years = vector(mode = "list", length = length(albums))
# 
# names(album_years) = albums
# 
# album_years["Taylor Swift"] = 2006
# album_years["Fearless"] = 2008
# album_years["Speak Now"] = 2010
# album_years["Red"] = 2012
# album_years["1989"] = 2014
# album_years["reputation"] = 2017
```

## Sentiment Analysis

```{r}
ts_afinn = ts_lyrics_tidy %>%
  inner_join(get_sentiments("afinn")) %>%
  group_by(index = album) %>%
  summarise(sentiment = sum(value)) %>%
  mutate(method = "AFINN")

ts_bing_and_nrc = bind_rows(ts_lyrics_tidy %>%
                              inner_join(get_sentiments("bing")) %>%
                              mutate(method = "Bing"),
                            ts_lyrics_tidy %>% 
                              inner_join(get_sentiments("nrc")) %>%
                                filter(sentiment %in% c("positive", "negative")) %>%
                              mutate(method = "NRC")) %>%
  count(method, index = album, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)

albums_in_order = c("Taylor Swift", "Fearless", "Speak Now", "Red", "1989", "reputation")

# order albums by release date
ts_all_lexicons = bind_rows(ts_afinn, ts_bing_and_nrc) %>%
  mutate(index, index = factor(index, levels = albums_in_order, ordered = TRUE)) %>%
  arrange(index)
```

```{r}
ts_all_lexicons %>%
  ggplot(aes(index, sentiment, fill = method)) +
  geom_col(show.legend = FALSE) + 
  facet_wrap(~method, ncol = 1, scales = "free_y") +
  ggtitle("Sentiment of Taylor Swift albums over time")
```

**NRC** agrees quite closely with **AFINN**. **Bing** gives quite different results.

Overall consensus is that *Red* is the most positive. *1989* is the least positive (in fact, it is the only one that had an overall negative sentiment using the **Bing** lexicon).

Let's drill down to the song level. We'll go album-by-album.

```{r}
get_album_lyrics = function(album_name) {
  ts_lyrics %>% filter(album == album_name) %>% return()
}

album_lyrics = lapply(albums_in_order, get_album_lyrics)
names(album_lyrics) = albums_in_order

album_lyrics_tidy = lapply(album_lyrics, unnest_tokens, word, lyric)
names(album_lyrics_tidy) = albums_in_order
```

```{r}
calc_sentiments_by_song = function(album_tidy) {
  sentiments_afinn = album_tidy %>%
    inner_join(get_sentiments("afinn")) %>%
    group_by(index = track_title) %>%
    summarise(sentiment = sum(value)) %>%
    mutate(method = "AFINN")
  
  sentiments_bing_and_nrc = bind_rows(album_tidy %>% 
                                        inner_join(get_sentiments("bing")) %>%
                                        mutate(method = "Bing"),
                                      album_tidy %>%
                                        inner_join(get_sentiments("nrc")) %>%
                                        mutate(method = "NRC")) %>%
    count(method, index = track_title, sentiment) %>%
    spread(sentiment, n, fill = 0) %>%
    mutate(sentiment = positive - negative)
  
  sentiments_all_lex = bind_rows(sentiments_afinn, sentiments_bing_and_nrc)
  
  sentiments_all_lex
}

plot_sentiments_by_song = function(sentiments_by_song) {
  sentiment_plot <- sentiments_by_song %>%
    ggplot(aes(index, sentiment, fill = method)) +
    geom_col(show.legend = FALSE) + 
    facet_wrap(~method, ncol = 1, scales = "free_y")
  
  sentiment_plot
}
```

```{r}
# album_sentiments_by_song = lapply(album_lyrics_tidy, calc_sentiments_by_song)
# sentiment_plots = lapply(album_sentiments_by_song, plot_sentiments_by_song)

# n_albums = length(albums_in_order)
# n_col = floor(sqrt(n_albums))
# 
# grid.arrange(grobs = sentiment_plots, ncol = n_col)

# album_lyrics_tidy[[6]] %>% calc_sentiments_by_song() %>% plot_sentiments_by_song()

# split_album = function(album, max_group_size) {
#   album_sections = album %>% 
#     group_by(track_n %/% max_group_size) %>%
#     nest %>%
#     pull(data)
#   
#   album_sections
# }
# 
# album_lyrics_tidy_split = lapply(album_lyrics_tidy, split_album, 3)

all_songs_sents = lapply(album_lyrics_tidy, calc_sentiments_by_song) %>% bind_rows()

plot_sent_measure = function(lex, all_sents) {
  all_songs_lex = all_sents %>% filter(method == lex)
  song_lex_sorted = sort(all_songs_lex$sentiment, index.return = TRUE)
  
  plot(1:length(song_lex_sorted$x), song_lex_sorted$x, main = lex)
}

lex_vec = c("AFINN", "Bing", "NRC")

# par(mfrow = c(3, 1))
lapply(lex_vec, plot_sent_measure, all_sents = all_songs_sents)
```

```{r}
hist_sent_measure = function(lex, all_sents) {
  all_songs_lex = all_sents %>% filter(method == lex)
  song_lex_sorted = sort(all_songs_lex$sentiment, index.return = TRUE)
  
  hist(song_lex_sorted$x, main = lex)
}

lapply(lex_vec, hist_sent_measure, all_sents = all_songs_sents)
```

In all cases, we see an S-shaped curve. But there's a gap between the "middle" and the most negative songs. 

```{r}
neg_songs_nrc = all_songs_sents %>% filter(method == "NRC", sentiment < -10)
neg_songs_afinn = all_songs_sents %>% filter(method == "AFINN", sentiment < -30)
neg_songs_bing = all_songs_sents %>% filter(method == "Bing", sentiment < -30)

very_neg_songs = bind_rows(neg_songs_nrc, neg_songs_afinn, neg_songs_bing) %>% select(index) %>% unique()
very_neg_songs
```

```{r}
bing_word_counts = ts_lyrics_tidy %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()

bing_word_counts %>%
  group_by(sentiment) %>%
  top_n(20) %>%
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(y = "Contribution to sentiment",
       x = NULL) +
  coord_flip()
# 
# nrc_joy = get_sentiments("nrc") %>%
#   filter(sentiment == "joy")
# 
# nrc_disgust = get_sentiments("nrc") %>%
#   filter(sentiment == "disgust")
# 
# ts_lyrics_tidy %>%
#   inner_join(nrc_disgust) %>%
#   count(word, sort = TRUE)
```

Should we add "shake" to `stop_words`?

```{r}
ts_lyrics_tidy %>%
  filter(word == "shake") %>%
  group_by(track_title) %>%
  count()
```

This is why "Shake it Off" is so overwhelmingly negative. But is it really a negative song?

But should we use absolute sentiment measures? Longer songs have more potential for extreme values (higher variance). So we can look at sentiment / length: a sort of "rate".

```{r}
song_word_counts = ts_lyrics_tidy %>%
  group_by(track_title) %>%
  count() %>%
  ungroup()

all_songs_sents_sorted = all_songs_sents %>% arrange(index)

( ( all_songs_sents_sorted %>% filter(method == "AFINN") %>% select(index) )== (song_word_counts$track_title) ) %>% sum() == length(unique(song_word_counts$track_title))
```

Alternatively, we do a join - does not require sorting.

```{r}
all_sents_lengths = 
  all_songs_sents %>% left_join(song_word_counts, by = c("index" = "track_title"))
```

```{r}
all_sents_lengths = all_sents_lengths %>% 
  mutate(rate = sentiment / n)

all_sents_lengths
```

```{r}
# plot(afinn_lengths$rate %>% sort())
```

```{r}
# each song should be a single row
# separate columns for each sentiment measure
all_sents_lengths %>% pivot_wider(names_from = method, values_from = sentiment) -> all_sents_lengths_pivoted

all_sents_lengths_pivoted
```

```{r}
# all_songs_sents %>% filter(method == "AFINN") %>% pivot_wider(names_from = method, values_from = sentiment) -> afinn_lengths
# all_songs_sents %>% filter(method == "Bing")  %>% pivot_wider(names_from = method, values_from = sentiment) -> bing_lengths
# all_songs_sents %>% filter(method == "NRC")  %>% pivot_wider(names_from = method, values_from = sentiment) -> nrc_lengths

all_sents_lengths %>% filter(method == "AFINN") %>% pivot_wider(names_from = method, values_from = rate) -> afinn_lengths
all_sents_lengths %>% filter(method == "Bing")  %>% pivot_wider(names_from = method, values_from = rate) -> bing_lengths
all_sents_lengths %>% filter(method == "NRC")  %>% pivot_wider(names_from = method, values_from = rate) -> nrc_lengths

afinn_lengths %>% left_join(bing_lengths, by = "index") %>% left_join(nrc_lengths, by = "index") -> tidy_sents
```

So instead of eyeball-clustering, let's *actually* cluster songs by these measures. We can assess the 

Hopefully a 3-cluster solution yields a breakdown similar to the one I've described: very negative, middle, very positive.

```{r}
# tidy_sents %>% select(AFINN, Bing, NRC) %>% kmeans(centers = 3) -> kclust_3

tidy_sents %>% select(AFINN, Bing, NRC) %>% scale() -> tidy_just_sents_stand

k_vals = 1:9

kclusts <- tibble(k = 1:9) %>%
  mutate(
    kclust = map(k, ~kmeans(tidy_just_sents_stand, .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, tidy_just_sents_stand)
  )

kclusts
```

```{r}
clusters <- kclusts %>%
  unnest(tidied)

assignments <- kclusts %>% 
  unnest(augmented)

clusterings <- kclusts %>%
  unnest(glanced, .drop = TRUE)
```

```{r}
ggplot(clusterings, aes(k, tot.withinss)) +
  geom_line()
```

There are a few possible "elbows", but I'd say any `k` from 3 to 5 is reasonable. (this comment was made using raw sentiment values)

Using "rates" makes `k` = 4 looks pretty good.

Standardized: `k` = 4 looks good.

```{r}
kclust_4 = assignments %>% filter(k == 4)
allSame <- function(x) length(unique(x)) == 1
allSame(kclust_4$kclust)
```

```{r}
kclust_4 %>% filter(kclust_4$kclust[[1]]$cluster == 3)
clusters
```


Let's try a Gaussian mixture.


```{r}
library(mclust)
```

```{r}
ts_gmm_BIC = mclustBIC(tidy_just_sents_stand)
summary(ts_gmm_BIC)
plot(ts_gmm_BIC)
```

GMM really prefers `k` = 2.

```{r}

```


The line of reasoning:

If an artist has more very-negative songs, then that artist is negative. (justify that this is a reasonable metric)

Taylor Swift has more very-negative songs. (perform computations)

Taylor Swift is a negative artist. (conclusion)

```{r}
library(reticulate)

n_songs = ts_lyrics %>% 
  select(track_title) %>% 
  unique() %>% 
  nrow()
```

Use clusters to define cutoffs (like we did above)

Get someone else's library

```{r}
beyonce_lyrics = read_csv("lyrics_beyonce.csv")
View(beyonce_lyrics)
```



