---
title: "Taylor's Reputation: Sentiment Analysis to Determine Selective Exposure to Information about Taylor Swift's Negativity"
author: "Carson Zhang (cxzhang4)"
date: "5/11/2020"
output:
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# code used to collect data from Genius
# artists and albums manually collected from Wikipedia

# library(genius)

# artist_albums <- tribble(
#  ~artist, ~album,
#  "Taylor Swift", "Taylor Swift",
#  "Taylor Swift", "Fearless",
#  "Taylor Swift", "Speak Now",
#  "Taylor Swift", "Red",
#  "Taylor Swift", "1989",
#  "Taylor Swift", "reputation",
#  "Taylor Swift", "Lover",
#  "Beyonce", "Dangerously in Love",
#  "Beyonce", "B'Day",
#  "Beyonce", "I Am... Sasha Fierce",
#  "Beyonce", "4",
#  "Beyonce", "Beyonce",
#  "Beyonce", "Lemonade",
#  "Britney Spears", "...Baby One More Time",
#  "Britney Spears", "Oops!... I Did It Again",
#  "Britney Spears", "Britney",
#  "Britney Spears", "In the Zone",
#  "Britney Spears", "Blackout",
#  "Britney Spears", "Circus",
#  "Britney Spears", "Femme Fatale",
#  "Britney Spears", "Britney Jean",
#  "Britney Spears", "Glory",
#  "Nicki Minaj", "Pink Friday",
#  "Nicki Minaj", "Pink Friday: Roman Reloaded",
#  "Nicki Minaj", "The Pinkprint",
#  "Nicki Minaj", "Queen",
#  "Adele", "19",
#  "Adele", "21",
#  "Adele", "25",
#  "Katy Perry", "Katy Hudson",
#  "Katy Perry", "One of the Boys",
#  "Katy Perry", "Teenage Dream",
#  "Katy Perry", "Prism",
#  "Katy Perry", "Witness",
#  "Lady Gaga", "The Fame Monster",
#  "Lady Gaga", "Born This Way",
#  "Lady Gaga", "Artpop",
#  "Lady Gaga", "Joanne",
#  "Lady Gaga", "Chromatica",
#  "Mariah Carey", "Mariah Carey",
#  "Mariah Carey", "Emotions",
#  "Mariah Carey", "Music Box",
#  "Mariah Carey", "Daydream",
#  "Mariah Carey", "Butterfly",
#  "Mariah Carey", "Rainbow",
#  "Mariah Carey", "Glitter",
#  "Mariah Carey", "Charmbracelet",
#  "Mariah Carey", "The Emancipation of Mimi",
#  "Mariah Carey", "E=MC2",
#  "Mariah Carey", "Memoirs of an Imperfect Angel",
#  "Mariah Carey", "Me. I Am Mariah... The Elusive Chanteuse",
#  "Mariah Carey", "Caution",
#  "Whitney Houston", "Whitney Houston",
#  "Whitney Houston", "Whitney",
#  "Whitney Houston", "I'm Your Baby Tonight",
#  "Whitney Houston", "My Love Is Your Love",
#  "Whitney Houston", "Just Whitney...",
#  "Whitney Houston", "I Look to You",
#  "Celine Dion", "Unison",
#  "Celine Dion", "Celine Dion",
#  "Celine Dion", "The Colour of My Love",
#  "Celine Dion", "Falling into You",
#  "Celine Dion", "Let's Talk About Love",
#  "Celine Dion", "These Are Special Times",
#  "Celine Dion", "A New Day Has Come",
#  "Celine Dion", "One Heart",
#  "Celine Dion", "Miracle",
#  "Celine Dion", "Taking Chances",
#  "Celine Dion", "Loved Me Back to Life",
#  "Celine Dion", "Courage",
#  "Madonna", "Madonna",
#  "Madonna", "Like a Virgin",
#  "Madonna", "True Blue",
#  "Madonna", "Like a Prayer",
#  "Madonna", "Erotica",
#  "Madonna", "Bedtime Stories",
#  "Madonna", "Ray of Light",
#  "Madonna", "Music",
#  "Madonna", "American Life",
#  "Madonna", "Confessions on a Dance Floor",
#  "Madonna", "Hard Candy",
#  "Madonna", "MDNA",
#  "Madonna", "Rebel Heart",
#  "Madonna", "Madame X",
 # "Rihanna", "Music of the Sun",
 # "Rihanna", "A Girl like Me",
 # "Rihanna", "Good Girl Gone Bad: Reloaded",
 # "Rihanna", "Rated R",
 # "Rihanna", "Loud",
 # "Rihanna", "Talk That Talk",
 # "Rihanna", "Unapologetic",
 # "Rihanna", "Anti"
# )

# artist_albums = tribble(
#         ~artist, ~album,
#         "Taylor Swift", "Lover"
# )


# female_pop_star_lyrics = artist_albums %>%
#  add_genius(artist, album)

# write_csv(female_pop_star_lyrics, "female_pop_star_lyrics.csv")
```

```{r}
library(tidyverse)
library(tidytext)
library(gridExtra)
library(broom)
```

```{r}
female_pop_star_lyrics = read_csv("data/female_pop_star_lyrics.csv")

# pre-processing: each row now contains a single word because
# we perform basic sentiment analysis at the word level
fps_lyrics_tidy = female_pop_star_lyrics %>% 
  unnest_tokens(word, lyric)
```

```{r}
# compute overall sentiment of each song using AFINN
fps_afinn = fps_lyrics_tidy %>%
  inner_join(get_sentiments("afinn")) %>%
  group_by(index = track_title) %>%
  summarise(sentiment = sum(value)) %>%
  mutate(method = "AFINN")

# compute overall sentiment of each song using Bing and NRC
fps_bing_and_nrc = bind_rows(fps_lyrics_tidy %>%
                              inner_join(get_sentiments("bing")) %>%
                              mutate(method = "Bing"),
                            fps_lyrics_tidy %>% 
                              inner_join(get_sentiments("nrc")) %>%
                                filter(sentiment %in% c("positive", "negative")) %>%
                              mutate(method = "NRC")) %>%
  count(method, index = track_title, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)

# combine all of the results
fps_all = bind_rows(fps_afinn, fps_bing_and_nrc)
```

```{r}
artists_albums = female_pop_star_lyrics %>% 
  select(artist, track_title) %>% distinct()

# add the album that each track comes from
fps_all_albums = fps_all %>% 
  inner_join(artists_albums, by = c("index" = "track_title"))

# transform the data: each song only has one row
# each sentiment value (AFINN, Bing, NRC) is its own variable (column)
afinn_pivoted = fps_all_albums %>% 
  filter(method == "AFINN") %>% 
  pivot_wider(names_from = method, values_from = sentiment)
bing_pivoted = fps_all_albums %>% 
  filter(method == "Bing") %>% 
  pivot_wider(names_from = method, values_from = sentiment)
nrc_pivoted = fps_all_albums %>% 
  filter(method == "NRC") %>% 
  pivot_wider(names_from = method, values_from = sentiment)

fps_pivoted = afinn_pivoted %>% 
  inner_join(bing_pivoted, by = c("index", "artist")) %>% 
  inner_join(nrc_pivoted, by = c("index", "artist"))
```

```{r}
fps_stand = fps_pivoted %>% 
  select(index, artist, AFINN, Bing, NRC)
```

```{r}
artist_song_counts = fps_stand %>% count(artist)
```

```{r}
# used to highlight a specific artist on a barplot
artist_highlight_bar = function() {
        list(geom_bar(stat = "identity"),
             coord_flip(),
             scale_fill_manual(values = c("yes" = "tomato", "no" = "gray"),
                               guide = FALSE)
        )
}

# used to plot proportions
prop_aes = aes(x = reorder(artist, prop), y = prop, fill = highlight)
# used to plot raw numbers
n_aes = aes(x = reorder(artist, n.x), y = n.x, fill = highlight)
```

```{r}
fps_totals = fps_stand %>% 
  group_by(artist) %>% 
  summarise(AFINN_total = sum(AFINN), Bing_total = sum(Bing), NRC_total = sum(NRC))

fps_props = fps_totals %>%
  inner_join(artist_song_counts, by = "artist") %>% 
  mutate(AFINN_rate = AFINN_total / n, 
         Bing_rate = Bing_total / n, 
         NRC_rate = NRC_total / n)
```

```{r}
# used to transform dataframe to highlight a specific artist when plotting
highlight_artist = function(df, artist_name) {
  df %>% 
    mutate(highlight = ifelse(artist == artist_name, "yes", "no")) %>%
    return()
}
```

## Introduction

## Method

### Data collection and preprocessing

We define the population that Taylor Swift comes from as popular female artists.

Specifically, an artist is sufficiently popular if they have sold over 200 million records. We analyze popular artists because we hypothesize that the belief arises from overfitting based on a small number of popular songs. Therefore, we want the artists in our population to be reasonable candidates for the same belief and belief-generating process. We choose 200 million as the cutoff because that is the lower bound of the group that Taylor Swift is in on Wikipedia's list of best-selling musical artists.

We choose to only analyze female artists because I believe that women are perceived differently from men. I claim tbat societal perceptions of momen's roles in romantic relationships are different from men's roles, so we restrict our analysis to women in order to control for these differences.

This results in 

Lyrics were collected from Genius using the “genius” R package. Genius was chosen as a data source because of its frequent collaborations with popular artists, suggesting that it contains at least some official lyrics, and not simply crowdsourced lyrics.

As collected, the dataset had one song line per row. Because we are performing sentiment analysis at the word level, tokens were unnested so that each row contained a single word.

Notably, no stop words were removed. This is because the existence of stop words is already accounted for by the sentiment dictionaries: if a word does not indicate a sentiment, it simply will not exist in the dictionary, and therefore will not affect the sentiment of the document.

###

```{r}
fps_props %>%
  highlight_artist("Taylor Swift") %>%
  ggplot(aes(x = reorder(artist, AFINN_rate), y = AFINN_rate, fill = highlight)) +
  artist_highlight_bar() +
  ggtitle("Artists ranked by AFINN (standardized)")

fps_props %>%
  highlight_artist("Taylor Swift") %>%
  ggplot(aes(x = reorder(artist, Bing_rate), y = Bing_rate, fill = highlight)) +
  artist_highlight_bar() +
  ggtitle("Artists ranked by Bing (standardized)")

fps_props %>%
  highlight_artist("Taylor Swift") %>%
  ggplot(aes(x = reorder(artist, NRC_rate), y = NRC_rate, fill = highlight)) +
  artist_highlight_bar() +
  ggtitle("Artists ranked by NRC (standardized)")
```

```{r}
fps_stand %>%
  ggplot(aes(AFINN, color = artist)) +
  geom_freqpoly(binwidth = 20) +
  ggtitle("AFINN distribution by artist")

fps_stand %>%
  ggplot(aes(Bing, color = artist)) +
  geom_freqpoly(binwidth = 20) +
  ggtitle("Bing distribution by artist")

fps_stand %>%
  ggplot(aes(NRC, color = artist)) +
  geom_freqpoly(binwidth = 20) +
  ggtitle("NRC distribution by artist")
```

```{r}
songs_less_point05 = fps_stand %>% 
  mutate(AFINN_cdf = cume_dist(AFINN), 
         Bing_cdf = cume_dist(Bing), 
         NRC_cdf = cume_dist(NRC)) %>%
  filter(AFINN_cdf < 0.05 & Bing_cdf < 0.05 & NRC_cdf < 0.05)
```

```{r}
songs_less_point1 = fps_stand %>% 
  mutate(AFINN_cdf = cume_dist(AFINN), Bing_cdf = cume_dist(Bing), NRC_cdf = cume_dist(NRC)) %>%
  filter(AFINN_cdf < 0.1 & Bing_cdf < 0.1 & NRC_cdf < 0.1)
```

```{r}
less_point05_with_prop = fps_stand %>%
  mutate(AFINN_cdf = cume_dist(AFINN), 
         Bing_cdf = cume_dist(Bing), 
         NRC_cdf = cume_dist(NRC)) %>%
  filter(AFINN_cdf < 0.05 & Bing_cdf < 0.05 & NRC_cdf < 0.05) %>%
  count(artist) %>%
  inner_join(artist_song_counts, by = "artist") %>%
  mutate(prop = n.x / n.y )%>% 
  mutate(highlight = ifelse(artist == "Taylor Swift", "yes", "no"))

less_point1_with_prop = fps_stand %>%
  mutate(AFINN_cdf = cume_dist(AFINN), 
         Bing_cdf = cume_dist(Bing), 
         NRC_cdf = cume_dist(NRC)) %>%
  filter(AFINN_cdf < 0.1 & Bing_cdf < 0.1 & NRC_cdf < 0.1) %>%
  count(artist) %>%
  inner_join(artist_song_counts, by = "artist") %>%
  mutate(prop = n.x / n.y )%>% 
  mutate(highlight = ifelse(artist == "Taylor Swift", "yes", "no"))
```

```{r}
less_point05_with_prop %>%
  ggplot(prop_aes) +
  artist_highlight_bar() +
  ggtitle("Artists ranked by proportion of very negative songs (top 10% in all dicts)")

less_point1_with_prop %>%
  ggplot(prop_aes) +
  artist_highlight_bar() +
  ggtitle("Artists ranked by proportion of very negative songs (top 10% in all dicts)")
```

## T-tests

```{r}
artist_names = fps_stand$artist %>% unique()
other_artists = artist_names[!(artist_names == "Taylor Swift")]
```

```{r}
ts_afinn = ( fps_stand %>% filter(artist == "Taylor Swift") )$AFINN
afinn_pvals = vector("numeric", length = length(other_artists))
i = 1
for (other_artist in other_artists) {
  other_artist_afinn = ( fps_stand %>% filter(artist == other_artist) )$AFINN
  
  test_result = t.test(ts_afinn, other_artist_afinn)
  print(test_result$p.value)
  
  afinn_pvals[i] = test_result$p.value
  i = i + 1
}
```

```{r}
ts_bing = ( fps_stand %>% filter(artist == "Taylor Swift") )$Bing
bing_pvals = vector("numeric", length = length(other_artists))
i = 1
for (other_artist in other_artists) {
  other_artist_bing = ( fps_stand %>% filter(artist == other_artist) )$Bing
  
  test_result = t.test(ts_bing, other_artist_bing)
  print(test_result$p.value)
  
  bing_pvals[i] = test_result$p.value
  i = i + 1
}
```

```{r}
ts_NRC = ( fps_stand %>% filter(artist == "Taylor Swift") )$NRC
NRC_pvals = vector("numeric", length = length(other_artists))
i = 1
for (other_artist in other_artists) {
  other_artist_NRC = ( fps_stand %>% filter(artist == other_artist) )$NRC
  
  test_result = t.test(ts_NRC, other_artist_NRC)
  print(test_result$p.value)
  
  NRC_pvals[i] = test_result$p.value
  i = i + 1
}
```

```{r}
hist(afinn_pvals)

afinn_sig_idx = afinn_pvals < 0.1

mean(afinn_sig_idx)

sum(afinn_sig_idx)
```

```{r}
hist(bing_pvals)

bing_sig_idx = bing_pvals < 0.1

mean(bing_sig_idx)

sum(bing_sig_idx)
```

```{r}
hist(NRC_pvals)

NRC_sig_idx = NRC_pvals < 0.1

mean(NRC_sig_idx)

sum(NRC_sig_idx)
```

```{r}
afinn_sig_idx
bing_sig_idx
NRC_sig_idx
```

```{r}
sent_pvals = data.frame(AFINN = afinn_pvals, 
                        Bing = bing_pvals, 
                        NRC = NRC_pvals) %>% 
  as_tibble()

sent_pvals = data.frame(AFINN = afinn_pvals, 
                        Bing = bing_pvals, 
                        NRC = NRC_pvals, 
                        row.names = other_artists)
```

```{r}
# print values without scientific notation
options(scipen = 99999)

sent_pvals
```

## cite

Wikipedia

Hart et al

